Nail: A practical interface generator for data formats

### 2014-12-03 paper/login/eval.tex
done with pass

\subsection{Real-World Formats}
\label{s:eval-formats}

@@

\begin{figure}[tb]
\centering
\begin{tabular}{lrl}
\toprule
\textbf{Protocol} & \textbf{LoC} & \textbf{Challenging features} \\ 
\midrule
DNS packets & 48+64 & Label compression,\\
  & & count fields \\
ZIP archives & 92+78 & Checksums, offsets, \\ 
  & & variable length trailer, \\
  & & compression \\
Ethernet  & 16+0\phantom{0} & --- \\
ARP       & 10+0\phantom{0} & --- \\
IP        & 25+0\phantom{0} & Total length field, options \\
UDP       &  7+0\phantom{0} & Checksum, length field \\
ICMP      &  5+0\phantom{0} & Checksum \\
\bottomrule
\end{tabular}

@@

\paragraph{DNS} In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
for DNS packets, shown in Figure~\ref{fig:dns-full}.
The grammar corresponds almost directly to the diagrams in RFC 1035,
which defines DNS\@. Nail's dependent fields handle DNS's count fields, and
transformations represent label compression. Both of these features are awkward at best to handle
with existing tools.

\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
by PKWare. At the end of each ZIP file is an
\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
located by scanning backwards from the end of the file until a magic number and a valid length field
is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
contains the magic number~\cite{wolf:berlinsides-zip}.
This header  contains points to the \emph{ZIP directory}, contains an entry every file
in the archive.  Each entry stores file metadata in addition to the
offset of a \emph{local file header}. The local file header duplicates
most information from the directory entry header and is followed immediately by the compressed
archive entry.
Duplicating information made sense when ZIP files were stored on floppy
disks with slow seek times and high fault rates, but nowadays lead to parsers being confused, such
as in the recent Android Master Key bug.

@@

Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
files transparently with transformations, and the programmer
is presented with a data structure containing the archive contents. Nail even allows parsing the
contents of an archive file -- allowing formats based on ZIP, such as Microsoft Office documents, to
be handled with one grammar.

@@

%code size unzip 6.0
%2821 lines
We implemented two applications---a
DNS server and an
\cc{unzip} program---based on the above grammars, and will compare the effort involved and the
resulting security to similar applications with hand-written parsers and with other parser generators.
We will use lines of code as a proxy for programmer effort. To evaluate security, we will argue how
our design avoids classes of vulnerabilities and fuzz-test one of our applications.
\begin{figure}[tb]
\centering
\smaller[0.5]
\begin{tabular}{lrr@{~}l}
\toprule
\textbf{Application}
  & \textbf{LoC w/ Nail}
  & \multicolumn{2}{c}{\textbf{LoC w/o Nail}} \\
\midrule
DNS server
  & 295
  & 683
  & (Hammer parser) \\

@@

We implemented a ZIP file extractor in 50 lines of C code, together with
92 lines of Nail grammar and 78 lines of C code implementing two stream
transforms (one for the DEFLATE compression algorithm with the help of
the zlib library, and one for finding the end-of-directory header).
 The \cc{unzip} utility contains a file \cc{extract.c}, which parses
ZIP metadata and calls various decompression routines in other files. This
file measures over 1,600 lines of C, which suggests that Nail is highly
effective at reducing manual input parsing code, even for the complex
ZIP file format.

In our full paper~\cite{bangert:nail}, we present a study of 15 ZIP parsing bugs.
11 of these vulnerabilities involved memory corruption during input handling, which generated code
is immune to by design. We also fuzz-tested our DNS server.
More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
parsers are not very reusable, as they build application-specific data structures and are tightly
coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
avoiding protocol misunderstandings.

>>>>

\subsection{Real-world formats}
\label{s:eval-formats}

@@

\begin{figure}[tb]
\smaller[0.5]
\centering
\begin{tabular}{lrl}
\toprule
\textbf{Protocol} & \textbf{LoC} & \textbf{Challenging features} \\ 
\midrule
DNS packets & 48+64 & Label compression,\\
  & & count fields \\
ZIP archives & 92+78 & Checksums, offsets, \\ 
  & & variable length trailer, \\
  & & compression \\
Ethernet  & 16+0\phantom{0} & --- \\
ARP       & 10+0\phantom{0} & --- \\
IP        & 25+0\phantom{0} & Total length field, options \\
UDP       &  7+0\phantom{0} & Checksum, length field \\
ICMP      &  5+0\phantom{0} & Checksum \\
\bottomrule
\end{tabular}

@@

\paragraph{DNS} In \S\ref{s:design}, we introduced Nail's syntax with a grammar
for DNS packets, shown in Figure~\ref{fig:dns-full}.
The grammar corresponds almost directly to the diagrams in RFC 1035,
which defines DNS\@. Nail's dependent fields handle DNS's count fields, and
transformations represent label compression. Both of these features are awkward at best to handle
with existing tools.

\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
by PKWare. At the end of each ZIP file is an
\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
located by scanning backwards from the end of the file until a magic number and a valid length field
is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
contains the magic number~\cite{wolf:berlinsides-zip}.
This header  contains pointers to the \emph{ZIP directory}, which contains an entry for every file
in the archive.  Each entry stores file metadata in addition to the
offset of a \emph{local file header}. The local file header duplicates
most information from the directory entry header and is followed immediately by the compressed
file data.
Duplicating information made sense when ZIP files were stored on floppy
disks with slow seek times and high fault rates, but nowadays lead to parsers being confused, such
as in the Android ZIP file vulnerability described in \S\ref{sec:intro}.

@@

Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
files transparently with transformations; the programmer
is presented with a data structure containing the archive contents. Nail even allows parsing the
contents of an archive file, allowing formats based on ZIP, such as Microsoft Office documents, to
be handled with one grammar.

@@

%code size unzip 6.0
%2821 lines
We implemented two applications---a
DNS server and an
\cc{unzip} program---based on the above grammars, and compare the effort involved and the
resulting security to similar applications with hand-written parsers and with other parser generators.
We use lines of code as a proxy for programmer effort. To evaluate security, we argue how
our design avoids classes of vulnerabilities and we also fuzz-test one of our applications.
\begin{figure}[tb]
\centering
\smaller[0.5]
\begin{tabular}{lrr@{~}l}
\toprule
\textbf{Application}
  & \textbf{LoC w/ Nail}
  & \multicolumn{2}{c}{\textbf{LoC w/o Nail}} \\
\midrule
DNS server
  & 295
  & 683
  & (Hammer parser) \\

@@

We implemented a ZIP file extractor in 50 lines of C code, together with
92 lines of Nail grammar and 78 lines of C code implementing two stream
transforms (one for the DEFLATE compression algorithm with the help of
the zlib library, and one for finding the end-of-directory header).

As a point of comparison, the existing Linux \cc{unzip} utility contains a file \cc{extract.c}, which parses
ZIP metadata and calls various decompression routines in other files. This
file measures over 1,600 lines of C, which suggests that Nail is highly
effective at reducing manual input parsing code, even for the complex
ZIP file format.

In our full paper~\cite{bangert:nail}, we present a study of 15 ZIP parsing bugs.
11 of these vulnerabilities involved memory corruption during input handling,
to which our generated code is immune by design.
We also fuzz-tested our DNS server, finding no bugs.
More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
parsers are not very reusable, as they build application-specific data structures and are tightly
coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
avoiding protocol misunderstandings.

### 2014-12-03 paper/login/intro.tex
done with pass

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past.
Most of these attacks involve memory corruption; therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

>>>>

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
\label{sec:intro}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past.
Most of these attacks involve memory corruption; therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.
